---
title: "Halictus_Wings_2022"
author: "Charlie Thrift"
date: '2022-08-28'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document, we analyze variation in wing morphologies between three species of _Halictus_ and two populations of _Halictus tripartitus_ using geometric morphometrics.

Necessary inputs for this script include:
(1) TPS files containing plotted landmark data for each wing photograph or "occurrence"
(2) CSV file containing all bee occurrence data, eg. all Hymenoptera from CCBER's GBIF or Symbiota page. This is used to match the TPS wing occurrence with its proper species and population information.

##Load Packages
```{r packages, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(ggbiplot)
library(geomorph) #NOTE: download XQuartz if using a Mac
library(vegan)
library(devtools)
library(stringr)
library(caret)
library(klaR)
```

##Load Data
###Part One: TPS data
The TPS data are in one file, which was created using tpsUtil32 and tps___(name of other program?)___. Here, we read these data and specify a few parameters for the readland.tps function. There are no curves to be read (readcurves=FALSE), and any negative values for landmarks are negative in Cartesian space (negNA=FALSE). If negNA were true, the function would read any negative landmarks as a missing landmark (this is because while landmarking, if a character is obscured you can "skip" landmarks within the landmarking software. We do not include any specimens missing any landmarks in this analysis).
```{r}
#Read in the TPS data for all bee wing images
all_tps_data <- readland.tps("TPS_files/AppendedTPS_all_02sep22.TPS",
                             specID = c("imageID"), readcurves = FALSE, 
                             warnmsg = TRUE,negNA = FALSE)
```

###Part Two: Bee data
Read bee data by first reading all Hymenoptera. After this chunk of code, all bee data is saved as __df1__ and is ready to be merged with the TPS data.
```{r}
#need to have "beedata26jun22.csv" saved in working directory
all_CCBER_hymenoptera <- read_csv("beedata26jun22.csv")

#clean up data. Only keep relevant columns
df1 <- all_CCBER_hymenoptera
df1 <- df1[,c("catalogNumber", "family", "genus", "scientificName",
                   "decimalLatitude", "decimalLongitude", "eventDate",
                   "day", "month", "year", "identifiedBy", "sex", "startDayOfYear")]
#clean up data: prepare to match on catalogNumber
df1 <- df1 %>% separate(catalogNumber, c("UCSB", "number")) #issue btwn _ vs - in catalogNumber
df1 <- df1 %>% tidyr::unite('catalogNumber', sep = "_", c('UCSB','number')) #solving it by switching to _ from - in the downloaded Hymenoptera data
```

##Generalized Procrustes Analaysis (GPA)
Run GPA on TPS data, to then generate coordinate data for each landmark on each specimen. In this code, we start with tps data and end with a CSV of coordinate data for each specimen. This can then be merged with the bee data generated above.
```{r}
all_tps_gpa <- gpagen(all_tps_data, PrinAxes = TRUE)

write.csv(all_tps_gpa$coords, "all_tps_02sep22.csv", row.names = TRUE)
write.csv(all_tps_gpa$Csize, "all_Csize_tps_02sep22.csv", row.names = TRUE)
```

##Merge Landmark Data with Bee Collection/Curation Data

###Adjust the Landmark Data to usable format
Read in the landmark data. Read in the Csize data for each specimen. Adjust each.
```{r}
lmdata <- read.csv("all_tps_02sep22.csv")
Csize <- read.csv("all_Csize_tps_02sep22.csv")

Csize1 <- Csize 
Csize2 <- setNames(cbind(rownames(Csize1), Csize1, row.names = NULL), 
         c("number_delete", "specimenID", "Csize")) #add row names
Csize3 <- Csize2[,c(2:3)] #delete the first column (which is just numbering)

df_transpose = t(lmdata) #transpose the landmark data
df_transpose1 <- df_transpose
df_transpose1 <- df_transpose1[c(2:961),]

#Format of LM data is currently two rows per specimen
#with one row being X coordinate values and one row
#being Y coordinate values. Below, we split into two 
#data frames and then stitch them back together to get
#18 different variables for the 9 landmarks
#(9 X coordinates and 9 Y coordinates)
lm1 <- df_transpose1
lmX <- lm1
lmY <- lm1
lmX1 <- lmX
lmX1 <- data.frame(lmX1)
lmX2 <- setNames(cbind(rownames(lmX1), lmX1, row.names = NULL),
                 c("name", "LM1x", "LM2x", "LM3x",
                   "LM4x", "LM5x", "LM6x", "LM7x", "LM8x", "LM9x"))
lmXonly <- lmX2[str_detect(lmX2$name, "X.UCSB"), ]
#repeat with Y
lmY1 <- lmY
lmY1 <- data.frame(lmY1)
lmY2 <- setNames(cbind(rownames(lmY1), lmY1, row.names = NULL),
                 c("name", "LM1y", "LM2y", "LM3y",
                   "LM4y", "LM5y", "LM6y", "LM7y", "LM8y", "LM9y"))
lmYonly <- lmY2[str_detect(lmY2$name, "Y.UCSB"), ]

lmXonly1 <- lmXonly
lmYonly1 <- lmYonly

lmXonly2 <- lmXonly1 %>% 
    tidyr::separate(name,                      
                    c("X","UCSB", "barcode", "wing","species", 
                      "location", "wingSide"), extra='drop') %>%
    tidyr::unite('catalogNumber', c('UCSB','barcode')) 
#drop any wings that were Right instead of Left
lmXonly3 <- lmXonly2[lmXonly2$wingSide %in% c("ed", NA), ] #remove any "right" wings
lmXonly4 <- lmXonly3[lmXonly3$species %in% c("edited", "far",
                "lig", "tri"), ] #remove any "right" wings

lmXonly5 <- lmXonly4[,c(2,7:15)]
##now: lmXonly5 has x coordinate values for all 9 landmarks, and just the catalogNumber
###repeat for Y
lmYonly2 <- lmYonly1 %>% 
    tidyr::separate(name,                      
                    c("Y","UCSB", "barcode", "wing","species", 
                      "location", "wingSide"), extra='drop') %>%
    tidyr::unite('catalogNumber', c('UCSB','barcode')) 
#drop any wings that were Right instead of Left
lmYonly3 <- lmYonly2[lmYonly2$wingSide %in% c("ed", NA), ] #remove any "right" wings
lmYonly4 <- lmYonly3[lmYonly3$species %in% c("edited", "far",
                "lig", "tri"), ] #remove any "right" wings

lmYonly5 <- lmYonly4[,c(2,7:15)]


########
#Now: unite the Y and X coordinate dataframes into just one
lm_both <- merge(lmXonly5, lmYonly5, by=c("catalogNumber"))
####Final step: add in the Csize column
Csize4 <- Csize3
Csize5 <- Csize4 %>% 
    tidyr::separate(specimenID,                      
                    c("UCSB", "barcode", "wing","species", 
                      "location", "wingSide"), extra='drop') %>%
    tidyr::unite('catalogNumber', c('UCSB','barcode')) 

Csize6 <- Csize5[,c(1,6)]

lm_both_size <- merge(lm_both, Csize6, by=c("catalogNumber"))
####Done with landmark data. "lm" has each specimen and 18 variables for lm coordinates
####plus 1 variable for Csize
```

###Merge Landmark and Curation Data
```{r}
lm_data <- merge(lm_both_size, df1, by=c("catalogNumber"))

####removing some specimens from the analysis
allsets_data123 <- lm_data
##Need to drop the images with missing/obscured landmarks
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00041576")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00041646")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037403")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037414")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00038716")
# drop: poor photo quality
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037491")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037465")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037432")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037425")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012006")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012259")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037414")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037403")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037397")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037386")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037356")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037336")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00036708")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00036688")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00035322")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00034772")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00030903")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00030503")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00030428")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00030379")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00030327")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00030289")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00030048")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00022211")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00021402")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00015171")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012463")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012424")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012342")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012277")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012056")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012049")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012026")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00012018")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00041576")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00036237")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037365")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00037381")
allsets_data123 <- subset(allsets_data123, catalogNumber != "UCSB_IZC00036773")
lm_data_clean <- allsets_data123
#remove 2 rows that have only Hymenoptera as their sci name
lm_data_clean <- subset(lm_data_clean, scientificName != "Hymenoptera")
```

##Summarizing Specimens
```{r}
lm <- lm_data_clean
lm$scientificName <- as.factor(lm$scientificName)
summary(lm$scientificName)

lm$species[lm$scientificName == 
             "Halictus tripartitus Cockerell, 1895"]<- "H. tripartitus"
lm$species[lm$scientificName == 
             "Halictus ligatus Say, 1837"]<- "H. ligatus"
lm$species[lm$scientificName == 
             "Halictus farinosus Smith, 1853"]<- "H. farinosus"
lm$species <- as.factor(lm$species)
summary(lm$species)

lm$location[lm$decimalLatitude <= 34.234 & 
                 lm$decimalLatitude >= 33.86 & 
                 lm$decimalLongitude >= -120.05 & 
                 lm$decimalLongitude <= -119.45]<- "Santa Cruz Island"
lm$location[lm$decimalLatitude <= 36 & 
                 lm$decimalLatitude >= 34.113887 & 
                 lm$decimalLongitude >= -121 & 
                 lm$decimalLongitude <= -116]<- "Mainland"
lm$location <- as.factor(lm$location)
summary(lm$location)

lm <- subset(lm, location != "NA")

ggplot(lm, aes(decimalLongitude, decimalLatitude)) +
       geom_point(aes(color=location))

## Check count of species and locations
table(lm$species, lm$location)

```
## Create subset with equal numbers of specimesn
```{r}
table(lm$species, lm$location)

lm_standard <- lm
lm_ligatus <- subset(lm_standard, species == "H. ligatus")
lm_tripartitus <- subset(lm_standard, species == "H. tripartitus")
lm_farinosus <- subset(lm_standard, species == "H. farinosus")

lm_ligatus_standard <- lm_ligatus[sample(nrow(lm_ligatus),size = 43), ]
lm_tripartitus_standard <- lm_tripartitus[sample(nrow(lm_tripartitus),size = 43), ]
lm_farinosus_standard <- lm_farinosus #because this is the lowest value already

lm_standard_species <- rbind(lm_ligatus_standard, 
                             lm_tripartitus_standard,
                             lm_farinosus_standard)


table(lm_standard_species$species, lm_standard_species$location)


#now make a data frame with equal numbers of just h tripartitus by location
table(lm_tripartitus$location)
tri_main <- subset(lm_tripartitus, location == "Mainland")
tri_island <- subset(lm_tripartitus, location == "Santa Cruz Island")

##based on table, need to reduce mainland down to 149, to match island
tri_main_standard <- tri_main[sample(nrow(tri_main),size = 149), ]
tri_island_standard <- tri_island #because this is the lowest value already
tri_data_standard <- rbind(tri_main_standard, tri_island_standard)
table(tri_data_standard$location)
```


## Run Principal Component Analysis (PCA)
Here, we run a principal component analysis on the coordinate data. 
```{r}
lm <- lm_standard_species #setting the LM data to be the one with the standardized
##number of specimens per each species

#Run PCA
lm.pca <- prcomp(lm[,c(2:19)], #with 2:19 we drop Csize
                       center = TRUE, scale. = TRUE)


tri <- tri_data_standard #setting the tripartitus LM data to be the one with standard
#number sizes per each location (149 each)

#Run PCA
tri.pca <- prcomp(tri[,c(2:19)], #with 2:19 we drop Csize
                       center = TRUE, scale. = TRUE)
```

## Run Statistical Tests
Here, we run the statistical tests for this project. First is the ANOSIM. This test uses the coordinate data as an input, and groups by species, to show if the groups are distinctly separate or not. ANOSIM test statistic close to 1 is highly different. A significance value lower than 0.05 is statistically significant.
#### ANOSIM for Species
```{r}
#Now, also remove the text columns, so the ANOSIM can run on just the numerical data
lm_anosim <- lm[, c(2:19)] #with 2:19 we drop Csize

#Run ANOSIM
##Here, run ANOSIM using Euclidean method
anosim(lm_anosim, 
       lm$species, 
       distance = "euclidean", ##EUCLIDEAN
       permutations = 999)
##Here, run ANOSIM using Manhattan method
#anosim(lm_anosim, 
#       lm$species, 
#       distance = "manhattan", ##MANHATTAN
#       permutations = 999)
```

### ANOSIM for Location
```{r}
#Now, remove the text columns, so the ANOSIM can run on just the numerical data
tri_anosim <- tri[, c(2:19)] #with 2:19 we drop Csize


#Run ANOSIM
##Here, run ANOSIM using Euclidean method
anosim(tri_anosim, 
       tri$location, 
       distance = "euclidean", ##EUCLIDEAN
       permutations = 999)
##Here, run ANOSIM using Manhattan method
anosim(tri_anosim, 
       tri$location, 
       distance = "manhattan", ##MANHATTAN
       permutations = 999)
```

## Visualize PCA on Biplots
```{r}
ggbiplot(lm.pca,
  ellipse = TRUE,  
  labels=rownames(lm$species), 
  var.axes = FALSE, 
  alpha = 0.75,
  groups=lm$species)+
  ggtitle("Halictus PCA (by species)")+ 
  theme_minimal()+
  theme(legend.position = "bottom")+
  scale_colour_manual(name="species", values= c("darkorchid",
                                                "chartreuse4",
                                                "red3"))
```

```{r}
ggbiplot(tri.pca,
  ellipse = TRUE,  
  labels=rownames(tri$location), 
  var.axes = FALSE, 
  alpha = 0.75,
  groups=tri$location)+
  ggtitle("H. tripartitus PCA (by location)")+ 
  theme_minimal()+
  theme(legend.position = "bottom")+
  scale_colour_manual(name="species", values= c("tan3", "cornflowerblue"))
```

## Cross Validation Tests
Here, see if the computer can accurately assign "unknown" specimens to the correct species, or, to the correct location within a single species.
```{r}

lm_cross <- lm
lm_cross <- lm_cross[, c(2:19, 33)] #use 2:19 to drop Csize
#species cross validation test
## Turn grouping column into a factor
lm_cross$species <- as.factor(lm_cross$species)
## Define training control
set.seed(123)
species_control <- trainControl(method = "repeatedcv", 
                              number = 10, repeats = 3)
# Train the model
species_model <- train(species ~., data = lm_cross, method = "nb",
               trControl = species_control)
# Summarize the results
print(species_model)
confusionMatrix(species_model)


####Cross Validation for Location
tri_cross <- tri
tri_cross <- tri_cross[,c(2:19, 34)] #use 2:19 to drop Csize
## Turn grouping column into a factor
tri_cross$location <- as.factor(tri_cross$location)
## Define training control
set.seed(123)
location_control <- trainControl(method = "repeatedcv", 
                              number = 10, repeats = 3)
# Train the model
location_model <- train(location ~., data = tri_cross, method = "nb",
               trControl = location_control)
# Summarize the results
print(location_model)
confusionMatrix(location_model)

```
## NMDS and Strength of Factors in LM data
```{r}
lm_nmds <- lm[ , c(2:19)] #use 2:19 to drop Csize
lm_nmds_in <- as.matrix(lm_nmds) #switch to matrix
lm_nmds_out <- metaMDS(lm_nmds_in, distance = "maha") #may want to pick a diff. distance matrix

env2 <- lm_nmds
en2 <- envfit(lm_nmds_out, env2, permutations = 999)
en2$vectors


## Repeat with just location in Tri
tri_nmds <- tri[ , c(2:19)] #use 2:19 to drop Csize
tri_nmds_in <- as.matrix(tri_nmds) #switch to matrix
tri_nmds_out <- metaMDS(tri_nmds_in, distance = "maha") #may want to pick a diff. distance matrix

env3 <- tri_nmds
en3 <- envfit(tri_nmds_out, env3, permutations = 999)
en3$vectors
```


## Make Tables for Mean Landmark Values
```{r}
main_mean <- tri_main_standard
island_mean <- tri_island_standard

tri_table <- data.frame(Landmark=(c("1x", "1y",
                                    "2x", "2y",
                                    "3x", "3y",
                                    "4x", "4y",
                                    "5x", "5y",
                                    "6x", "6y",
                                    "7x", "7y",
                                    "8x", "8y",
                                    "9x", "9y")),
                        Rsquared=(c(en3[["vectors"]][["r"]][["LM1x"]],
                                   en3[["vectors"]][["r"]][["LM1y"]],
                                   en3[["vectors"]][["r"]][["LM2x"]],
                                   en3[["vectors"]][["r"]][["LM2y"]],
                                   en3[["vectors"]][["r"]][["LM3x"]],
                                   en3[["vectors"]][["r"]][["LM3y"]],
                                   en3[["vectors"]][["r"]][["LM4x"]],
                                   en3[["vectors"]][["r"]][["LM4y"]],
                                   en3[["vectors"]][["r"]][["LM5x"]],
                                   en3[["vectors"]][["r"]][["LM5y"]],
                                   en3[["vectors"]][["r"]][["LM6x"]],
                                   en3[["vectors"]][["r"]][["LM6y"]],
                                   en3[["vectors"]][["r"]][["LM7x"]],
                                   en3[["vectors"]][["r"]][["LM7y"]],
                                   en3[["vectors"]][["r"]][["LM8x"]],
                                   en3[["vectors"]][["r"]][["LM8y"]],
                                   en3[["vectors"]][["r"]][["LM9x"]],
                                   en3[["vectors"]][["r"]][["LM9y"]])),
                        P_value=(c(en3[["vectors"]][["pvals"]][["LM1x"]],
                                   en3[["vectors"]][["pvals"]][["LM1y"]],
                                   en3[["vectors"]][["pvals"]][["LM2x"]],
                                   en3[["vectors"]][["pvals"]][["LM2y"]],
                                   en3[["vectors"]][["pvals"]][["LM3x"]],
                                   en3[["vectors"]][["pvals"]][["LM3y"]],
                                   en3[["vectors"]][["pvals"]][["LM4x"]],
                                   en3[["vectors"]][["pvals"]][["LM4y"]],
                                   en3[["vectors"]][["pvals"]][["LM5x"]],
                                   en3[["vectors"]][["pvals"]][["LM5y"]],
                                   en3[["vectors"]][["pvals"]][["LM6x"]],
                                   en3[["vectors"]][["pvals"]][["LM6y"]],
                                   en3[["vectors"]][["pvals"]][["LM7x"]],
                                   en3[["vectors"]][["pvals"]][["LM7y"]],
                                   en3[["vectors"]][["pvals"]][["LM8x"]],
                                   en3[["vectors"]][["pvals"]][["LM8y"]],
                                   en3[["vectors"]][["pvals"]][["LM9x"]],
                                   en3[["vectors"]][["pvals"]][["LM9y"]])),
                        Mainland=(c(mean(main_mean$LM1x),
                                 mean(main_mean$LM1y),
                                 mean(main_mean$LM2x),
                                 mean(main_mean$LM2y),
                                 mean(main_mean$LM3x),
                                 mean(main_mean$LM3y),
                                 mean(main_mean$LM4x),
                                 mean(main_mean$LM4y),
                                 mean(main_mean$LM5x),
                                 mean(main_mean$LM5y),
                                 mean(main_mean$LM6x),
                                 mean(main_mean$LM6y),
                                 mean(main_mean$LM7x),
                                 mean(main_mean$LM7y),
                                 mean(main_mean$LM8x),
                                 mean(main_mean$LM8y),
                                 mean(main_mean$LM9x),
                                 mean(main_mean$LM9y))),
                        Island=(c(mean(island_mean$LM1x),
                                 mean(island_mean$LM1y),
                                 mean(island_mean$LM2x),
                                 mean(island_mean$LM2y),
                                 mean(island_mean$LM3x),
                                 mean(island_mean$LM3y),
                                 mean(island_mean$LM4x),
                                 mean(island_mean$LM4y),
                                 mean(island_mean$LM5x),
                                 mean(island_mean$LM5y),
                                 mean(island_mean$LM6x),
                                 mean(island_mean$LM6y),
                                 mean(island_mean$LM7x),
                                 mean(island_mean$LM7y),
                                 mean(island_mean$LM8x),
                                 mean(island_mean$LM8y),
                                 mean(island_mean$LM9x),
                                 mean(island_mean$LM9y))))

tri_table

tri_mean <- lm_tripartitus_standard
lig_mean <- lm_ligatus_standard
far_mean <- lm_farinosus_standard

species_table <- data.frame("Landmark"=(c("1x", "1y",
                                    "2x", "2y",
                                    "3x", "3y",
                                    "4x", "4y",
                                    "5x", "5y",
                                    "6x", "6y",
                                    "7x", "7y",
                                    "8x", "8y",
                                    "9x", "9y")),
                            Rsquared=(c(en2[["vectors"]][["r"]][["LM1x"]],
                                   en2[["vectors"]][["r"]][["LM1y"]],
                                   en2[["vectors"]][["r"]][["LM2x"]],
                                   en2[["vectors"]][["r"]][["LM2y"]],
                                   en2[["vectors"]][["r"]][["LM3x"]],
                                   en2[["vectors"]][["r"]][["LM3y"]],
                                   en2[["vectors"]][["r"]][["LM4x"]],
                                   en2[["vectors"]][["r"]][["LM4y"]],
                                   en2[["vectors"]][["r"]][["LM5x"]],
                                   en2[["vectors"]][["r"]][["LM5y"]],
                                   en2[["vectors"]][["r"]][["LM6x"]],
                                   en2[["vectors"]][["r"]][["LM6y"]],
                                   en2[["vectors"]][["r"]][["LM7x"]],
                                   en2[["vectors"]][["r"]][["LM7y"]],
                                   en2[["vectors"]][["r"]][["LM8x"]],
                                   en2[["vectors"]][["r"]][["LM8y"]],
                                   en2[["vectors"]][["r"]][["LM9x"]],
                                   en2[["vectors"]][["r"]][["LM9y"]])),
                            P_value=(c(en2[["vectors"]][["pvals"]][["LM1x"]],
                                   en2[["vectors"]][["pvals"]][["LM1y"]],
                                   en2[["vectors"]][["pvals"]][["LM2x"]],
                                   en2[["vectors"]][["pvals"]][["LM2y"]],
                                   en2[["vectors"]][["pvals"]][["LM3x"]],
                                   en2[["vectors"]][["pvals"]][["LM3y"]],
                                   en2[["vectors"]][["pvals"]][["LM4x"]],
                                   en2[["vectors"]][["pvals"]][["LM4y"]],
                                   en2[["vectors"]][["pvals"]][["LM5x"]],
                                   en2[["vectors"]][["pvals"]][["LM5y"]],
                                   en2[["vectors"]][["pvals"]][["LM6x"]],
                                   en2[["vectors"]][["pvals"]][["LM6y"]],
                                   en2[["vectors"]][["pvals"]][["LM7x"]],
                                   en2[["vectors"]][["pvals"]][["LM7y"]],
                                   en2[["vectors"]][["pvals"]][["LM8x"]],
                                   en2[["vectors"]][["pvals"]][["LM8y"]],
                                   en2[["vectors"]][["pvals"]][["LM9x"]],
                                   en2[["vectors"]][["pvals"]][["LM9y"]])),
                        "Halictus tripartitus"=(c(mean(tri_mean$LM1x),
                                 mean(tri_mean$LM1y),
                                 mean(tri_mean$LM2x),
                                 mean(tri_mean$LM2y),
                                 mean(tri_mean$LM3x),
                                 mean(tri_mean$LM3y),
                                 mean(tri_mean$LM4x),
                                 mean(tri_mean$LM4y),
                                 mean(tri_mean$LM5x),
                                 mean(tri_mean$LM5y),
                                 mean(tri_mean$LM6x),
                                 mean(tri_mean$LM6y),
                                 mean(tri_mean$LM7x),
                                 mean(tri_mean$LM7y),
                                 mean(tri_mean$LM8x),
                                 mean(tri_mean$LM8y),
                                 mean(tri_mean$LM9x),
                                 mean(tri_mean$LM9y))),
                        "Halictus ligatus"=(c(mean(lig_mean$LM1x),
                                 mean(lig_mean$LM1y),
                                 mean(lig_mean$LM2x),
                                 mean(lig_mean$LM2y),
                                 mean(lig_mean$LM3x),
                                 mean(lig_mean$LM3y),
                                 mean(lig_mean$LM4x),
                                 mean(lig_mean$LM4y),
                                 mean(lig_mean$LM5x),
                                 mean(lig_mean$LM5y),
                                 mean(lig_mean$LM6x),
                                 mean(lig_mean$LM6y),
                                 mean(lig_mean$LM7x),
                                 mean(lig_mean$LM7y),
                                 mean(lig_mean$LM8x),
                                 mean(lig_mean$LM8y),
                                 mean(lig_mean$LM9x),
                                 mean(lig_mean$LM9y))),
                         "Halictus farinosus"=(c(mean(far_mean$LM1x),
                                 mean(far_mean$LM1y),
                                 mean(far_mean$LM2x),
                                 mean(far_mean$LM2y),
                                 mean(far_mean$LM3x),
                                 mean(far_mean$LM3y),
                                 mean(far_mean$LM4x),
                                 mean(far_mean$LM4y),
                                 mean(far_mean$LM5x),
                                 mean(far_mean$LM5y),
                                 mean(far_mean$LM6x),
                                 mean(far_mean$LM6y),
                                 mean(far_mean$LM7x),
                                 mean(far_mean$LM7y),
                                 mean(far_mean$LM8x),
                                 mean(far_mean$LM8y),
                                 mean(far_mean$LM9x),
                                 mean(far_mean$LM9y))))
species_table
```


